---
title: "Figures"
output: pdf_document
classoption: landscape
fontsize: 12pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(dev = 'pdf')
pdf.options(encoding='ISOLatin2')
# knitr::opts_chunk$set(root.dir = normalizePath('..'))
# knitr::opts_knit$set(root.dir = normalizePath('..'))

```



```{r load, include=FALSE}
# INTRO -------------------------------------------------------------------
rm(list = ls())
graphics.off()
devtools::install_github('matthew-phelps/CholeraDataDK', force = F)

library(tidyverse)
library(CholeraDataDK)
library(scales) # required for some ggplot functions
library(cowplot)
source("functions.R") # required for rotating text in ggplot 
source("plots/plotHelperFunctions.R")

# LOAD --------------------------------------------------------------------

load("data/data-viz-prep.Rdata")
load("data/r0.Rdata")
all_cases_temp <- cholera_daily_data
mort <- cph_mort_rates_10yr
pop <- cph_pop1853_10yr
attack <- cph_age_attack_rate
cases <- cholera_daily_data_towns
cen <- census
dead_all_cause <- all_cause_1852_1854
rownames(cen) <- NULL

# Re-order levels of factor so that plotting works: http://goo.gl/CD2fEC
age_char <- as.character(mort$age_range)
mort$age_range <- factor(age_char, levels = c(age_char))

```

```{r season, fig.height=4, fig.width=8, message=FALSE}

options(warn = -1)
all_cases <- all_cases_temp %>%
  arrange(city, date) %>%
  filter(city != "brandholm")
# To plot all epidemics on same calendar year - put them on a "dummy" year
all_cases$season <- paste("100",
                          format(all_cases$date, "%m"),
                          format(all_cases$date, "%d"),
                          sep = "-")
broad_st$season <-  paste("100",
                          format(broad_st$date, "%m"),
                          format(broad_st$date, "%d"),
                          sep = "-")
all_cases$season <- as.Date(all_cases$season)
broad_st$season <- as.Date(broad_st$season)

lab_x <- min(all_cases$season)


plot_season <- all_cases %>%
  SeasonPlot(lab_x = lab_x, lab_size = 4, txt_size = 12, line_size = 0.8,
             broad_street = TRUE, broad_street_data = broad_st)
plot_season
```

## Figure 1   
Daily cholera case incidence per 10,000 people and seasonality of outbreaks in three cities with inset map showing city locations and populations at time of outbreak. Mortality counts from the Broad Street outbreak in London (1854) are included in grey as reference.



```{r allcause, fig.height=4, fig.width=8}
# # To long format data
# mort_gender <- gather(mort[, c("age_range", "male_mort2", "female_mort2")],
#                       key = gender,
#                       value = mort_rate,
#                       c(male_mort2, female_mort2))

# 1 - ALL_CAUSE MORTALITY  ---------------------------------

# To shade 1853, need a rectangle that covers only this year for all y
all_cause <- gather(dead_all_cause, cause, dead, 4:6)
yrng <- range(0:4000)
start_53 <- as.Date("1853-06-01") # start of shading
end_53 <- as.Date("1853-10-01") # End of shading



all_cause_plot <- all_cause %>%
  deathsBasePlot(l_size = 0.7)%>%
  deathPoints(p_size = 2, space = FALSE) %>%
  deathsStyles(txt_size = 12)
all_cause_plot
```


## Figure 2  
Monthly number of recorded deaths in Copenhagen due to cholera, diarrhea, and all-causes from January 1852 to December 1854.
\newpage


```{r r0,, fig.height=5, fig.width=7}

# Remove time-dependent
r0 <- r0[r0$method != "TD", ]
R0 <- r0 %>%
  r0Plot(pd = 0.4) %>%
  r0PlotStyle(txt_size = 12)
R0
```

## Figure 3  
Comparing the reproductive number (R0) in three Danish locations using two different methods to estimate R0.
\newpage


```{r mortANDattack, fig.height=4.5, fig.width=8}
# DF with cholera mort as % of total mort & cholera mort rates in 1 df
limits <- aes(ymax = upper95, ymin = lower95,
              x= age_range, color = plot_var2)
dodge <- position_dodge(width=- 0.3)
dodge2 <- position_dodge(width=- 0.5)


# In order to makes space between cities in legend, need to plot empty "dummy" variables that will be white space. Creat empty factors:
chol_burden$plot_var2 <- chol_burden$plot_var
chol_burden$plot_var2 <- factor(chol_burden$plot_var2,
                                levels = c("korsoer.total_attack", 
                                           "korsoer.total_mort_rate",
                                           "a",
                                           "aalborg.total_attack",
                                           "aalborg.total_mort_rate",
                                           "b",
                                           "cph.total_attack",
                                           "cph.total_mort_rate"))

plot_chol_mort <- cholMortPlot(p_size = 2, line_size = 0.7) %>%
  mortPlotStyle(txt_size = 11)
plot_chol_mort
```

## Figure 4  
Cholera morbidity and mortality per 100 people disaggregated by age for three cities.
\newpage

```{r choleraDeaths, fig.height=4.5, fig.width=8}
limits = aes(ymax = hi95, ymin = low95, x = age_range)



chol_proportion_death_plot <- cholProportionDeaths() %>%
  proportionDeathStyle(txt_size = 12)
chol_proportion_death_plot


```

## Figure 5  
Proportion of all deaths in Copenhagen, 1853 attributed to cholera disaggregated by age.
\newpage



```{r quarterPanel, fig.height=5.5, fig.width=9}

cph_quarter_for_plot <- cph_quarters %>%
  filter(quarter != "Fra skibe" & quarter != "Ladegaarden" &
           quarter != "Vesterbro" & quarter != "Noerrebro" & quarter != "Oesterbro")
quarters <- quarterPanelIncidence(cph_quarter_for_plot)
quarters

```


## Figure S1  
Incidence rate per 100 people in the neighborhoods of Copenhagen, 1853.
\newpage


```{r RelativeRisks, fig.height=7, fig.width=8}

title <- "Relative risk of cholera morbidity by age & gender \n"
notes <- "*male is reference group"

# Attack RR
plot_attack <- rr_sic %>%
  rrAgeGender(pd = 0.4, pd2 = 0.6,
              line_size = 0.5) %>%
  rrAgeGenderStyle(txt_size = 12,  ylabel = "infection")



plot_mort <- rr_mrt %>%
  rrAgeGender(pd = 0.4, pd2 = 0.6,
              line_size = 0.5) %>%
  rrAgeGenderStyle(txt_size = 12,
                   ylabel = "death",
                   x_title = TRUE)

# MULTIPLOT 
plot_grid(plot_attack, plot_mort, nrow = 2, labels = c("A", "B"))
```

## Figure S2  
Relative risk of cholera (A) morbidity and (B) mortality by age and gender. Values above 1.0 indicates a greater risk in females as compared to males.



