---
output: pdf_document
classoption: landscape
fontsize: 10pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(dev = 'pdf')
pdf.options(encoding='ISOLatin2')
# knitr::opts_chunk$set(root.dir = normalizePath('..'))
# knitr::opts_knit$set(root.dir = normalizePath('..'))

```

\pagenumbering{gobble}

```{r load, include=FALSE}
# INTRO -------------------------------------------------------------------
rm(list = ls())
graphics.off()
devtools::install_github('matthew-phelps/CholeraDataDK', force = F)

library(tidyverse)
library(CholeraDataDK)
library(scales) # required for some ggplot functions
library(cowplot)
source("functions.R") # required for rotating text in ggplot 
source("plots/plotHelperFunctions.R")

# LOAD --------------------------------------------------------------------

load("data/data-viz-prep.Rdata")
load("data/r0.Rdata")
all_cases_temp <- cholera_daily_data
mort <- cph_mort_rates_10yr
pop <- cph_pop1853_10yr
attack <- cph_age_attack_rate
cases <- cholera_daily_data_towns
cen <- census
dead_all_cause <- all_cause_1852_1854
rownames(cen) <- NULL

# Re-order levels of factor so that plotting works: http://goo.gl/CD2fEC
age_char <- as.character(mort$age_range)
mort$age_range <- factor(age_char, levels = c(age_char))

```

```{r season, fig.height=4, fig.width=8, message=FALSE}

options(warn = -1)
all_cases <- all_cases_temp %>%
  arrange(city, date) %>%
  filter(city != "brandholm")
# To plot all epidemics on same calendar year - put them on a "dummy" year
all_cases$season <- paste("100",
                          format(all_cases$date, "%m"),
                          format(all_cases$date, "%d"),
                          sep = "-")
broad_st$season <-  paste("100",
                          format(broad_st$date, "%m"),
                          format(broad_st$date, "%d"),
                          sep = "-")
all_cases$season <- as.Date(all_cases$season)
broad_st$season <- as.Date(broad_st$season)

lab_x <- min(all_cases$season)


plot_season <- all_cases %>%
  SeasonPlot(lab_x = lab_x, lab_size = 4, txt_size = 12, line_size = 0.8,
             broad_street = TRUE, broad_street_data = broad_st)
plot_season
```




```{r allcause, fig.height=4, fig.width=8}
# # To long format data
# mort_gender <- gather(mort[, c("age_range", "male_mort2", "female_mort2")],
#                       key = gender,
#                       value = mort_rate,
#                       c(male_mort2, female_mort2))

# 1 - ALL_CAUSE MORTALITY  ---------------------------------

# To shade 1853, need a rectangle that covers only this year for all y
all_cause <- gather(dead_all_cause, cause, dead, 4:6)
yrng <- range(0:4000)
start_53 <- as.Date("1853-06-01") # start of shading
end_53 <- as.Date("1853-10-01") # End of shading



all_cause_plot <- all_cause %>%
  deathsBasePlot(l_size = 0.7)%>%
  deathPoints(p_size = 1.5, space = FALSE) %>%
  deathsStyles(txt_size = 12)
all_cause_plot
```




```{r r0,, fig.height=5, fig.width=7}

# Remove time-dependent
r0 <- r0[r0$method != "TD", ]
R0 <- r0 %>%
  r0Plot(pd = 0.4, line_size = 0.7) %>%
  r0PlotStyle(txt_size = 12)
R0
```



```{r mortANDattack, fig.height=4.5, fig.width=8}
# DF with cholera mort as % of total mort & cholera mort rates in 1 df
limits <- aes(ymax = upper95, ymin = lower95,
              x= age_range, color = plot_var2)
dodge <- position_dodge(width=- 0.3)
dodge2 <- position_dodge(width=- 0.5)


# In order to makes space between cities in legend, need to plot empty "dummy" variables that will be white space. Creat empty factors:
chol_burden$plot_var2 <- chol_burden$plot_var
chol_burden$plot_var2 <- factor(chol_burden$plot_var2,
                                levels = c("korsoer.total_attack", 
                                           "korsoer.total_mort_rate",
                                           "a",
                                           "aalborg.total_attack",
                                           "aalborg.total_mort_rate",
                                           "b",
                                           "cph.total_attack",
                                           "cph.total_mort_rate"))

plot_chol_mort <- cholMortPlot(p_size = 1.8, line_size = 0.7) %>%
  mortPlotStyle(txt_size = 11)
plot_chol_mort
```



```{r choleraDeaths, fig.height=4.5, fig.width=8}
limits = aes(ymax = hi95, ymin = low95, x = age_range)



chol_proportion_death_plot <- cholProportionDeaths(line_size = 1) %>%
  proportionDeathStyle(txt_size = 12)
chol_proportion_death_plot


```

\newpage


**Figure 1: Daily cholera case incidence and seasonality of outbreaks in three cities**  
Mortality counts from the Broad Street outbreak in London (1854) are included in grey as reference.  

**Figure 2: Monthly number of recorded deaths in Copenhagen due to cholera, diarrhea, and all-causes from January 1852 to December 1854**  

**Figure 3: The reproductive number ($R_{0}$) in three Danish cities**  

**Figure 4: Cholera morbidity and mortality disaggregated by age for three Danish cities**  

**Figure 5: Proportion of all deaths in Copenhagen, 1853, attributed to cholera disaggregated by age**  

**Figure 6: Transmission chains and modes**  
Solid lines indicate known transmission paths. Dotted lines indicate all possible transmission paths if one single path was not determinable
 
\newpage


**Table 1: Summary statistics**  

**Table 2: Socio-economic status and disease outcomes by neighborhood**  

**Table 3: Comparison of key  metrics in cholera outbreaks in historic and contempoary settings**  


\newpage


```{r quarterPanel, fig.height=5.5, fig.width=9}

cph_quarter_for_plot <- cph_quarters %>%
  filter(quarter != "Fra skibe" & quarter != "Ladegaarden" &
           quarter != "Vesterbro" & quarter != "Noerrebro" & quarter != "Oesterbro")
quarters <- quarterPanelIncidence(cph_quarter_for_plot)
quarters

```




```{r RelativeRisks, fig.height=7, fig.width=8}

title <- "Relative risk of cholera morbidity by age & gender \n"
notes <- "*male is reference group"

# Attack RR
plot_attack <- rr_sic %>%
  rrAgeGender(pd = 0.4, pd2 = 0.6,
              line_size = 0.5) %>%
  rrAgeGenderStyle(txt_size = 12,  ylabel = "infection")



plot_mort <- rr_mrt %>%
  rrAgeGender(pd = 0.4, pd2 = 0.6,
              line_size = 0.5) %>%
  rrAgeGenderStyle(txt_size = 12,
                   ylabel = "death",
                   x_title = TRUE)

# MULTIPLOT 
plot_grid(plot_attack, plot_mort, nrow = 2, labels = c("A", "B"))
```
\newpage





**Supplementary figure 1: Incidence rate across the neighborhoods of Copenhagen, 1853**  

**Supplementary figure 2: Relative risk of (A) cholera morbidity and (B) mortality by age and gender**  
Values above 1.0 indicates a greater risk in females as compared to males.  

**Supplementary table 1: Index cases and transmission chains around Denmark**  


